---
title: "Clear cell sarcoma - surgical treatment"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
author: 'Piotr Błoński'
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    toc_depth: 2
---
<style>
body {
text-align: justify}
</style>

```{r echo=FALSE, cache=FALSE, warning=FALSE, message=FALSE}
library(knitr)
opts_chunk$set(
  echo = FALSE,
  cache = FALSE,
  message = FALSE,
  warning = FALSE
)
# 
```


```{r setup}
library(survival)
library(lubridate)
library(ggsurvfit)
library(gtsummary)
library(tidycmprsk)
library(dplyr)
library(stringi)
library(purrr)
library(xddd)
library(survminer)
library(swimplot)

db <- read.csv("C:/Users/pbpol/OneDrive/Dokumenty/NIO/robocze/medstat/data/epsarc.csv",
               header = TRUE, sep = ',')
db <- db[which(db$x==2),]
db[db %in% c(99,999,9999)] <- NA

db <- data.frame(
  x = as.numeric(db$x),
  PID = as.numeric(db$PID),
  BIRTH = db$birth_date,
  SEX = as.numeric(db$sex),
  DIAG = db$diag_date,
  DEAD = as.numeric(db$dead_tf),
  DEATH_DATE = db$death_date,
  MARGIN = db$col85,
  R0 = as.numeric(db$col85 == 0),
  SITE = as.numeric(db$col34),
  PRESENTATION = as.character(db$col48),
  DIAG = db$diag_date,
  INITIAL_SURG = db$surg_prim_date,
  relapse_initial_tf = as.numeric(db$relapse_tf),
  relapse_initial_date = db$relapse_date,
  initial_R = db$surg_prim_R,
  ok = db$col46,
  surg_prim_nio = db$surg_prim_nio,
  SIZE_PRIM = as.numeric(db$col37.1),
  SIZE_REC = as.numeric(db$col36.3),
  TNM_N = as.numeric(db$col45),
  TNM_M = as.numeric(db$col46),
  PERIOPERATIVE_REGIMEN = as.numeric(db$col51))

db_surg <- read.csv("C:/Users/pbpol/OneDrive/Dokumenty/NIO/robocze/medstat/data/ccs_surg.csv",
                    header = TRUE, sep = ',')

db <- merge(db_surg, db, by='PID', all.x=T, all.y=F)
db <- db[which(db$ok == 0),]
db <- db[which(!duplicated(db$PID)),]
# db <- db[which(!(db$surgery_type2)%in%c('scar excision + SLNB', 'scar excision')),]

db <- db %>% mutate(
  relapse_date = ifelse(relapse_tf, relapse_date, DEATH_DATE),
  relapse_tf = ifelse(relapse_tf, 1, DEAD),
  relapse_initial_date = ifelse(relapse_initial_tf, relapse_initial_date, DEATH_DATE),
  relapse_initial_tf = ifelse(relapse_initial_tf, 1, DEAD),
  initial_R0 = as.numeric(initial_R==0),
  OS_MONTHS = as.duration(surgery_date %--% DEATH_DATE)/dmonths(1),
  RFS_MONTHS = as.duration(surgery_date %--% relapse_date)/dmonths(1),
  AGE = round(as.duration(BIRTH %--% surgery_date)/dyears(1),0),
  OS_SINCE_DIAG_MONTHS = as.duration(DIAG %--% DEATH_DATE)/dmonths(1),
  RFS_INITIAL_MONTHS = as.duration(INITIAL_SURG %--% relapse_initial_date)/dmonths(1),
  OS_INITIAL_MONTHS = as.duration(INITIAL_SURG %--% DEATH_DATE)/dmonths(1),
  TUMOR_SIZE = ifelse(as.numeric(surg_prim_nio), SIZE_PRIM, SIZE_REC),
  relapse_pattern = ifelse(((relapse_pattern=='No recurrence')&(DEAD==1)), 'Death without relapse', relapse_pattern),
  trt_type = recode(surg_prim_nio, '0'='Recurrence/Scar excision', '1'='Primary tumor')
)

db <- cbind(data.frame(ID=paste(rep('ID_', times=nrow(db)), 1:nrow(db)),sep=''), db)


ccs_rfs <- survfit2(Surv(RFS_MONTHS, relapse_tf)~1, data=db)
ccs_rfs_initial <- survfit2(Surv(RFS_INITIAL_MONTHS, relapse_initial_tf)~1, data=db)
ccs_os <- survfit2(Surv(OS_MONTHS, DEAD)~1, data=db)
ccs_os_since_diag <- survfit2(Surv(OS_SINCE_DIAG_MONTHS, DEAD)~1, data=db)

plt_rfs <- prettyplot_for_one(ccs_rfs, survivaltitle=c('Relapse-free', 'RFS'),
                              plotxlab='Years since the surgery in NIO', x_intervals = 60,
                              survivaltitle_suffix = 'since the NIO surgery',
                              plotcol = 'black',vertlines = c(24, 60),x_max = 12*20,
                              annotateshift = 60)
plt_os <- prettyplot_for_one(ccs_os, survivaltitle=c('Overall', 'OS'),
                              plotxlab='Years since the surgery', x_intervals = 60,
                              plotcol = '#5d28fc',vertlines = c(24, 60),x_max = 12*20,
                              annotateshift = 60, survivaltitle_suffix = 'since the NIO surgery')

plt_nio_rfs <- pretty_plot_for2(db$RFS_MONTHS, db$relapse_tf, db$surg_prim_nio,
                              curve1name = 'For the recurrent tumor/Scar radicalization',
                              curve2name = 'For the primary disease',
                              survivaltitle = c('Relapse-free', 'RFS'),
                              plotxlab = 'Years since the surgery',
                              legend_tit = 'Reason for the \n surgery at NIO:',
                              x_max = 12*15, x_intervals = 60,curve1col = 'orange',
                              curve2col = 'grey', add_pvalue = FALSE,
                              survivaltitle_suffix = 'since the NIO surgery')
plt_nio_os <- pretty_plot_for2(db$OS_MONTHS, db$DEAD, db$surg_prim_nio,
                              curve1name = 'For the recurrent tumor/Scar radicalization',
                              curve2name = 'For the primary disease',
                              survivaltitle = c('Overall', 'OS'),
                              plotxlab = 'Years since the surgery',
                              legend_tit = 'Reason for the \n surgery at NIO:',
                              x_max = 12*15, x_intervals = 60, curve1col = 'purple',
                              curve2col = 'black', add_pvalue = FALSE,
                              survivaltitle_suffix = 'since the NIO surgery')
plt_nio_os_initial <- pretty_plot_for2(db$OS_INITIAL_MONTHS, db$DEAD, db$surg_prim_nio,
                              curve1name = 'For the recurrent tumor/Scar radicalization',
                              curve2name = 'For the primary disease',
                              survivaltitle = c('Overall', 'OS'),
                              plotxlab = 'Years since the surgery',
                              legend_tit = 'Reason for the \n surgery at NIO:',
                              x_max = 12*15, x_intervals = 60, curve1col = 'khaki',
                              curve2col = 'darkblue', add_pvalue = FALSE,
                              survivaltitle_suffix = 'since the initial surgery')

dbswim <- db[,c('ID', 'RFS_MONTHS', 'relapse_pattern', 'surg_prim_nio', 'OS_MONTHS', 'DEAD')]
dbswim <- dbswim %>% mutate(
  surg_prim_nio = recode(surg_prim_nio, '0'='Recurrent tumor/Scar radicalization','1' = 'Primary tumor'),
  DEAD = recode(DEAD, '0'='Censored', '1'='Death'),
  OS_MONTHS = OS_MONTHS*dmonths(1)/dyears(1),
  RFS_MONTHS = RFS_MONTHS*dmonths(1)/dyears(1)
)
dbswim <- dbswim %>% rename('Relapse_pattern'='relapse_pattern', 'Status'='DEAD')

  
swim_plt<- swimmer_plot_piotra(dbswim, id='ID', end = 'RFS_MONTHS', name_fill = 'Relapse_pattern', stratify =  'surg_prim_nio',legend.title = element_text('Pattern of recurrence'))+
  swimmer_points(dbswim, id='ID', time = 'OS_MONTHS', name_shape = 'Status')+
  swimmer_lines(dbswim, id = 'ID', start = 'RFS_MONTHS', end = 'OS_MONTHS')+
  scale_y_continuous(name = 'Years since the surgery', breaks = seq(0, 18, by=3))
```

# Baseline characteristics {-}

```{r table1}
table1 <- rbind(tb(rep('', nrow(db)), 'Total number of patients'),
                tb(recode(db$SEX,'1'='Female','2'='Male'), 'Sex'),
                tbcon(db$AGE, 'Age [years]'),
                tb(recode(db$SITE,'1'='Back','2'='Chest','3'='Arm','4'='Forearm','5'='Tigh','6'='Lower leg','7'='Head & Neck','8'='Palm/feet','9'='Pelvis','10'='Abdomen','11'='Uterus'), 'Tumor site'),
                tb(recode(db$PRESENTATION,'0'='Primary tumor before biopsy','1'='Primary tumor after biopsy','2'='Scar after primary excision','3'='Recurrent tumor','4'='Nodal metastases (Only)','5'='Distant metastases','7'='Recurrent tumor'),'Patient status'),
                tb(recode(db$TNM_N, '1'='Yes', '0'='No'), 'Involvement of lymph nodes (at admission)'),
                tb(recode(db$TNM_M, '1'='Yes', '0'='No'), 'Distant metastases (at admission)')
                
)
colnames(table1) <- c('variable', 'n (%)')
kable(table1, row.names = FALSE)
```

# Types of surgery {-}

```{r}
table2 <- tb(db$surgery_type2, 'Type of surgery at NIO')
colnames(table2) <- c('', 'n (%)')
kable(table2, row.names = F)
```

# Peri-operative treatment {-}

```{r tabl3}
tab3 <- rbind(
  tb(recode(db$PERIOPERATIVE_REGIMEN, '0'= 'Surgery only',
'1'= 'CHTH - surgery',
'2'= 'RTH - surgery',
'3'= 'CHTH + RTH - surgery',
'4'= 'Surgery - CHTH',
'5'= 'Surgery - RTH',
'6'= 'Surgery - RTH - CHTH',
'7'= 'RTH - surgery - RTH',
'8'= 'RTH + CHTH - surgery - CHTH',
'9'= 'RTH + CHTH - surgery - RTH',
'10'= 'CHTH - surgery - CHTH',
'11'= 'RTH - surgery - CTH',
'12'= 'palliative'), 'Perioperative treatment modalities')
)
colnames(tab3) <- c('', 'n (%)')
kable(tab3, row.names = FALSE)
```

# Swimmer plot {-}


The RFS since the surgery is represented by the thick bars, whereas the thin lines show the survival after the relapse. The patients have been split into groups: treated for the primary tumor (upper box) or for recurrent tumor/scar radicalization after resection with inadequate margins (lower box).

```{r, fig.asp=0.55, fig.width=9}
swim_plt
```

# Survival since the surgery at NIO {-}

The median relapse-free survival (RFS) since the surgery at NIO was `r surv_median(ccs_rfs)$median %>% round(1)` months (95% CI: `r surv_median(ccs_rfs)$lower %>% round(1)` - `r surv_median(ccs_rfs)$upper %>% round(1)`) for all patients. 

The median overall survival since the surgery at NIO was `r surv_median(ccs_os)$median %>% round(1)` months (`r surv_median(ccs_os)$lower %>% round(1)` - `r surv_median(ccs_os)$upper %>% round(1)`) for all patients. 

```{r, fig.asp=0.55}
plt_rfs
plt_os
```


# Treatment at NIO for the primary disease vs for recurrent/incompletely resected {-}

```{r}
primary <- sum(db$surgery_type2 %in% c('primary tumor resection', 'primary tumor resection + TLND', 'primary tumor resection + SLNB'))
recurrent <- sum(db$surgery_type2 %in% c('local recurrence resection', 'local recurrence resection + TLND', 'local recurrence resection + SLNB', 'electrochemotherapy'))
radicalization <- sum(db$surgery_type2 %in% c('scar excision', 'scar excision + SLNB'))
```

Out of `r nrow(db)` patients who underwent the curative-intent surgery for clear-cell sarcoma, `r primary` patients (`r round(100*primary/nrow(db),1)`%) were treated for the primary disease, another `r recurrent` (`r round(100*recurrent/nrow(db),)`%) were treated for recurrent disease (relapse after primary treatment in regional center) and in `r radicalization` (`r round(100*radicalization/nrow(db),)`%) patients, there was performed a radicalization of a scar after the preceding resection with inadequate margins in the regional center.



```{r, fig.asp=0.55}
plt_nio_rfs
plt_nio_os
plt_nio_os_initial
```

